Solar wind predictions for the Parker Solar Probe orbit

### 2017-08-24 paperMVVB/sections/abstract.tex
velocity treatment in discussion and summary

\abstract
{In view of the planned near-Sun spacecraft mission Parker~Solar~Probe (PSP) (formerly Solar~Probe~Plus) the solar wind environment for its prime mission duration (2018--2025) and down to its intended closest perihelion (\num{9.86}~solar radii) is extrapolated using in situ data. The PSP mission will be humanity's first in situ exploration of the solar corona. Visiting this yet uncharted region is of special interest, because it will help answer hitherto unresolved questions on the heating of the solar corona and the source and acceleration of the solar wind and solar energetic particles. The solar wind extrapolation of this study is performed within the project Coronagraphic German And US Solar~Probe Survey (CGAUSS) which is the German contribution to the PSP mission as part of the Wide field Imager for Solar PRobe (WISPR).}	%context
{We present an empirical solar wind model for the inner heliosphere which is derived from Helios and OMNI in situ data. The German-US space probes Helios~1 and Helios~2 flew in the 1970s and observed solar wind in the ecliptic within heliocentric distances of \SIrange{0.29}{0.98}{\au}. The OMNI database consists of multi-spacecraft intercalibrated in situ data obtained near \SI{1}{\au}. The solar wind model is used together with sunspot number predictions to estimate the frequency distributions of major solar wind parameters PSP will encounter during its mission.}	%aims
{The model covers the solar wind's magnetic field strength and its plasma parameters proton velocity, density and temperature.
Their individual frequency distributions are represented with lognormal functions. In addition, we also consider the velocity distribution's bi-componental shape, consisting of a slower and a faster part. The model accounts for solar activity and for solar distance dependency by shifting of these lognormal distributions. We compile functional relations to solar activity by correlating and fitting the frequency distributions with the sunspot number (SSN), using almost five solar cycles of OMNI data. Further, based on the combined data set from both Helios probes, the parameters' frequency distributions are fitted with respect to solar distance to obtain power law dependencies. Finally, by combining the found solar cycle and solar distance relations, we obtain a simple dynamical solar wind model for the inner heliosphere, confined to the ecliptic region.}	%methods
{The inclusion of SSN predictions and the extrapolation to the PSP perihelion region enables us to estimate the solar wind environment for PSP's planned trajectory during its mission duration. The estimated solar wind median values during PSP's first perihelion are \SI{87}{\nT}, \SI{340}{\km\per\s}, \SI{4015}{\per\cm\cubed} and \SI{503000}{\K}. The modeled values for PSP’s closest perihelia are \SI{943}{\nT}, \SI{290}{\km\per\s}, \SI{9733}{\per\cm\cubed} and \SI{1930000}{\K}, where these velocity and temperature values are clearly overestimated in comparison with existing observations.}	%results
{This empirical model shows that solar wind acceleration and heating processes below about 20~solar radii limit a simple back-extrapolation from existing in situ measurements.}	%conclusions

>>>>

\abstract
{In view of the planned near-Sun spacecraft mission Parker~Solar~Probe (PSP) (formerly Solar~Probe~Plus) the solar wind environment for its prime mission duration (2018--2025) and down to its intended closest perihelion (\num{9.86}~solar radii) is extrapolated using in situ data. The PSP mission will be humanity's first in situ exploration of the solar corona. Visiting this yet uncharted region is of special interest, because it will help answer hitherto unresolved questions on the heating of the solar corona and the source and acceleration of the solar wind and solar energetic particles. The solar wind extrapolation of this study is performed within the project Coronagraphic German And US Solar~Probe Survey (CGAUSS) which is the German contribution to the PSP mission as part of the Wide field Imager for Solar PRobe (WISPR).}	%context
{We present an empirical solar wind model for the inner heliosphere which is derived from Helios and OMNI in situ data. The German-US space probes Helios~1 and Helios~2 flew in the 1970s and observed solar wind in the ecliptic within heliocentric distances of \SIrange{0.29}{0.98}{\au}. The OMNI database consists of multi-spacecraft intercalibrated in situ data obtained near \SI{1}{\au}. The solar wind model is used together with sunspot number predictions to estimate the frequency distributions of major solar wind parameters PSP will encounter during its mission.}	%aims
{The model covers the solar wind's magnetic field strength and its plasma parameters proton velocity, density and temperature.
Their individual frequency distributions are represented with lognormal functions. In addition, we also consider the velocity distribution's bi-componental shape, consisting of a slower and a faster part. The model accounts for solar activity and for solar distance dependency by shifting of these lognormal distributions. We compile functional relations to solar activity by correlating and fitting the frequency distributions with the sunspot number (SSN), using almost five solar cycles of OMNI data. Further, based on the combined data set from both Helios probes, the parameters' frequency distributions are fitted with respect to solar distance to obtain power law dependencies. Finally, by combining the found solar cycle and solar distance relations, we obtain a simple dynamical solar wind model for the inner heliosphere, confined to the ecliptic region.}	%methods
{The inclusion of SSN predictions and the extrapolation to the PSP perihelion region enables us to estimate the solar wind environment for PSP's planned trajectory during its mission duration. The estimated solar wind median values during PSP's first perihelion are \SI{87}{\nT}, \SI{340}{\km\per\s}, \SI{4015}{\per\cm\cubed} and \SI{503000}{\K}. The modeled values for PSP’s closest perihelia are \SI{943}{\nT}, \SI{290}{\km\per\s}, \SI{9733}{\per\cm\cubed} and \SI{1930000}{\K}, where these velocity and temperature values are clearly overestimated in comparison with existing observations. This empirical model shows that solar wind acceleration and heating processes below about 20~solar radii limit a simple back-extrapolation from existing in situ measurements.}	%results
{}	%conclusions

### 2017-08-24 paperMVVB/sections/content.tex
velocity treatment in discussion and summary

{\color{red} discuss high value zoom figures; read in Veselovsky2010}

@@

The frequency distribution data for the four solar wind parameters with respect to the radial distance from the Sun are plotted in Fig.~\ref{fig:mixed_fit_fixed_4_paper_f_plot}, together with their power law lognormal fits and the double lognormal fit for the velocity with their median values.
\begin{figure*}
	\includegraphics[width=18cm]{figures/mixed_fit_fixed_4_paper_f_plot.pdf}
	\caption{Frequency distributions of the four solar wind parameters with respect to the solar distance. Plotted are the binned Helios data and the power law lognormal fit models with their median values (white lines). The double lognormal model is used for the velocity, its slow and fast parts are indicated by dotted lines.}
	\label{fig:mixed_fit_fixed_4_paper_f_plot}
\end{figure*}
The model’s magnetic field strength is broader around values of \SI{40}{nT} at the lower distance boundary than the data's frequency distribution implies. This behavior is expected because of the distance independent shape approximation applied. The velocity and temperature models’ upper values show a higher abundance than the actual data, see also zoom boxes in Figs.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot} and \ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}.

@@

We extrapolate the derived empirical solar wind model (\ref{eq:general_sw_model}) to PSP’s orbital distance range and compare the results with those from the existing models shown in Fig.~\ref{fig:sw_extrapolation_ssn_d_plot}.
\begin{figure*}
	\includegraphics[width=18cm]{figures/sw_extrapolation_ssn_d_plot.pdf}
	\caption{Radial extrapolation of the solar wind median parameters to the PSP orbit region. The from Helios and OMNI measurements obtained models are extrapolated to the PSP region---for the extreme cases of solar minimum (SSN = 0) and maximum (SSN = 200). Note that there is a time lag to the SSN depending on the solar wind parameter. Also plotted are the radial dependencies of the analytic DQCS magnetic field model for solar minimum from \citet{Banaszkiewicz1998}, the slow wind velocity models from \citet{Wang2000} and \citet{Sheeley1997}, the density model from \citet{Leblanc1998} and the range of temperature measurements from \citet{Billings1959} and \citet{Liebenberg1975}.}
	\label{fig:sw_extrapolation_ssn_d_plot}
\end{figure*}

@@

%velocity:
The average velocity is found to decrease from \SI{340}{\km\per\s} at \SI{0.25}{\au} to about \SI{290}{\km\per\s} at \SI{0.046}{\au} for a SSN of 0. Whereas using a SSN of 200 it decreases from \SI{390}{\km\per\s} to \SI{330}{\km\per\s}. Comparing the results with the measurements by \citet{Sheeley1997} and \citet{Wang2000} shows an overestimation in our extrapolated slow solar wind velocity values for distances below about \SI{20}{\Rs}. They used LASCO coronagraph observations to track moving coronal features (blobs) in the distance range \SIrange{2}{30}{\Rs} to determine speed profiles and sources of the slow solar wind and they derived temperature and sonic point values for slow solar wind with the isothermal expansion model from \citet{Parker1958}. Therefore, it generally can be expected that PSP will encounter a slower solar wind environment close to the Sun than our model estimates and thus PSP will measure solar wind acceleration processes \hl{below} \SI{30}{\Rs} \citep{Sheeley1997,McComas2008}.

@@

\section{Discussion}
\label{sec:discussion}

We started the development of the empirical solar wind environment model for the near-ecliptic PSP orbit by lognormally fitting the about 40~years of in situ near-Earth solar wind data collected in the OMNI database, using the frequency distributions of the key solar wind parameters magnetic field strength, velocity, density and temperature. The OMNI multi-spacecraft solar wind data is intercalibrated and covers almost five solar cycles. It thus represents solar wind gathered at different phases of solar activity in the ecliptic plane. In the next step we investigated the yearly variation of the solar wind distribution functions along with the SSN over 53~years and derived linear dependencies of the solar wind parameters with the SSN. The radial dependencies of the solar wind distribution functions were then analyzed using Helios~1 and 2 data for the distance range \SIrange{0.29}{0.98}{\au} in bins of \SI{0.01}{\au}. The herefrom derived power law fit functions were used to scale the prior calculated SSN-dependent \SI{1}{\au} distribution fit functions to the PSP orbit, combined with SSN predictions for the years 2018--2025, i.e., for the prime mission phase that gets PSP to closest perihelia of 9.86~solar radii distance. The reason for performing the analysis this way is based on the fact that the OMNI solar wind database is much larger than the Helios database. Though, it is clear that the calculated distribution functions just represent first order estimates of the real solar wind to be encountered by PSP. The solar wind environment to be encountered will depend at times of PSP on the structure of the solar corona and underlying photospheric magnetic field and on the evolution and interaction of individual solar wind streams and superimposed CMEs and shocks.

However, the derived results are in good agreement with existing studies about near-Sun solar wind magnetic field strenghts and densities as shown in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}. The from direct measurements differing velocity and temperature extrapolations below about 20~solar radii indicate that PSP will indeed dive into the acceleration and heating regions of the solar wind to be expected at these distances (see Fig.~\ref{fig:sw_extrapolation_ssn_d_plot}). The near-Sun solar wind velocity at PSP perihelion is also expected to be slower than our model estimates, because the region of the Alfvénic critical surface, up to which the solar wind is believed to be accelerated, is predicted to lie in average around \SI{17}{\Rs} \citep[e.g.,][]{Sittler1999,Exarhos2000} and correlates with solar activity between \SI{15}{\Rs} in cycle minimum and \SI{30}{\Rs} in maximum \citep{Katsikas2010,Goelzer2014}.	% in average around 17~Rs (e.g., 0.6~Rs Schatten 1969, 17~Rs Sittler1999, 17~Rs Exarhos2000) and scales with solar activity between 15~Rs in minimum and 30~Rs in maximum (19~Rs Katsikas2010, 15–30~Rs Goelzer2014).

In our study we have not specifically investigated the occurrences of extreme solar wind parameters caused by CMEs or enhanced values due to stream interaction or co-rotating interaction regions. The Helios solar wind measurements plotted over radial distance in Fig.~\ref{fig:radial_fit_4_thesis_light_b_skip} show several extreme values far above the usual solar wind velocities, likely to be associated with individual CMEs. The results by \citet{Sachdeva2017} indicate that due to the prevailing of solar wind drag the speeds of fast CMEs will commonly slow down substantially from early distances of a few solar radii. Therefore, it is expected that PSP will encounter CMEs with much higher speeds than those observed during the Helios mission. Also, the magnetic field, density and temperature values are expected to be much larger than in the average solar wind in individual fast shock associated CME events. PSP will thus also substantially improve our understanding of the near-Sun evolution of CMEs and their expansion with radial distance.\\

{\color{red}
note special velocity treatment with double lognormal fits\\
slow/fast streams basically maintain characteristic speeds with solar cycle\\

derive heliocentric distance depending error...\\
error estimation for general model and extreme value tendencies\\

magnetic field distribution's with distance increasing high value tail -> source are compression regions (why with density no increase?); look into Parker1958's B-field formula...\\
varying shape with distance is indicator for internal physical processes (mixing/turbulence...)\\

slow/fast ratio SSN dependency\\
individual velocity part discussion -> there is no specific velocity threshold between slow and fast solar wind types, the velocity ranges of both types overlap.\\
Not only the slowest wind but also the fastest wind is expected to converge to the average speed (Sanchez-Diaz2016 p.~2835, using MHD-model -> very slow solar wind is continuation of slow wind) (because of interaction).\\
The ratio of both varies with solar activity, e.g., 3~years after maximum, polar coronal holes are observed to often have equatorial extensions (cite?). see and use \citet{Bougeret1984} p.~498...\\
for the overlapping velocity model the SSN dependency is steeper than for a simple threshold\\
}

\section{Summary}
For the near-Earth solar wind OMNI data and the Helios~1 and 2 data, obtained over the distance range \SIrange{0.29}{0.98}{\au}, we derived lognormal representations of the frequency distributions’ shapes of the four key solar wind parameters magnetic field strength, proton velocity, proton density and proton temperature. The dependencies of these frequency distributions on the solar activity cycle and on radial distance to the Sun have been modeled with analytical relations and extrapolated to the Parker Solar Probe orbit, taking into account predictions of the sunspot number. This empirical CGAUSS solar wind model for PSP, representing the solar wind’s solar activity and distance behavior, yields estimations for the solar wind median values at PSP’s first perihelion in 2018 and at PSP’s closest perihelion in 2024. The model yields the following main results for the bulk solar wind parameters:
\begin{itemize}
	\item The empirical dependency of the magnetic field strength median value on solar activity and radial distance is
	\begin{align}
		B_\text{med}(ssn,r) = (\SI{0.0131}{\nT} \cdot ssn + \SI{4.29}{\nT}) \cdot r^{-1.66}	\,.
	\end{align}
	This approximation appears to be valid for PSP's full distance range, however near its closest perihelion the actual median values might lie within \SI{20}{\%} higher. The estimated median values for PSP's first and closest perihelion are \SI{87}{\nT} and \SI{943}{\nT}.
	\item Averaging over fast and slow solar wind streams, the derived median velocity value dependence is
	\begin{align}
		v_\text{med}(ssn,r) = \left(\SI{1.31}{\km\per\s} \cdot ssn + \SI{429}{\km\per\s}\right) \cdot r^{0.099}	\,.
	\end{align}
	This relation is valid above about \SI{20}{\Rs}, below this solar distance it overestimates the actual velocities obtained from remote measurements. The estimated median values for PSP's first and closest perihelion are \SI{340}{\km\per\s} and \SI{290}{\km\per\s}.
	\item The density relation is found to be
	\begin{align}
		n_\text{med}(ssn,r) = \left(\SI{0.0038}{\per\cm\cubed} \cdot ssn + \SI{4.50}{\per\cm\cubed}\right) \cdot r^{-2.11}	\,.
	\end{align}
	It seems valid throughout the full PSP orbital distance range, even down to about \SI{8}{\Rs}. The estimated median values for PSP's first and closest perihelion are \SI{4015}{\per\cm\cubed} and \SI{9733}{\per\cm\cubed}.
	\item The derived correlation function for the temperature is
	\begin{align}
		T_\text{med}(ssn,r) = (\SI{197}{\K} \cdot ssn + \SI{57300}{\K}) \cdot r^{-1.10}	\,.
	\end{align}
	Near PSP's perihelion the validity of this relation is difficult to evaluate. The estimated median values for PSP's first and closest perihelion are \SI{503000}{\K} and \SI{1930000}{\K}.
\end{itemize}

{\color{red}
The velocity and temperature values are overestimated below \SI{20}{\Rs} in comparison with existing observations. This shows that solar wind acceleration and heating processes below \SI{20}{\Rs} limit the simple back-extrapolation from existing in situ measurements.\\
}

>>>>

%{\color{red} discuss high value zoom figures; read in Veselovsky2010}

@@

The frequency distribution data for the four solar wind parameters with respect to the radial distance from the Sun are plotted in Fig.~\ref{fig:mixed_fit_fixed_4_paper_f_plot}, together with their power law lognormal fits and the double lognormal fit for the velocity with their median values.
\begin{figure*}
	\includegraphics[width=18cm]{figures/mixed_fit_fixed_4_paper_f_plot.pdf}
	\caption{Frequency distributions of the four solar wind parameters with respect to the solar distance. Plotted are the binned Helios data and the power law lognormal fit models with their median values (white lines). The double lognormal model is used for the velocity, its slow and fast parts are indicated by dotted lines.}
	\label{fig:mixed_fit_fixed_4_paper_f_plot}
\end{figure*}
The model’s magnetic field strength is broader around values of \SI{40}{nT} at the lower distance boundary than the data's frequency distribution implies. This behavior is expected because of the distance independent shape approximation applied. The velocity and temperature models’ upper values generally show a higher abundance than the actual data, see also zoom boxes in Figs.~\ref{fig:histogram_fits_4_a_zoom_paper_pdfplot} and \ref{fig:histogram_fits_V_a_zoom_dbl_paper_pdfplot}. The velocity model's with distance increasing high velocity tail comes from using the same exponent for both slow and fast components. This effect is not seen in the data, more specifically, not only the slowest wind but also the fastest wind is expected to converge to more average speeds \citep{Sanchez-Diaz2016}.
%\citep{Sanchez-Diaz2016} p.~2835, using MHD-model -> very slow solar wind is continuation of slow wind) (because of interaction)

@@

We extrapolate the derived empirical solar wind model (\ref{eq:general_sw_model}) to PSP’s orbital distance range and compare the results with those from the existing models shown in Fig.~\ref{fig:sw_extrapolation_ssn_d_plot}.
\begin{figure*}
	\includegraphics[width=18cm]{figures/sw_extrapolation_ssn_d_plot.pdf}
	\caption{Radial extrapolation of the solar wind median parameters to the PSP orbit region. The from Helios and OMNI measurements obtained models are extrapolated to the PSP region---for the range between the extreme cases of solar minimum and maximum, i.e., $0 \le ssn \le 200$ (shaded areas). Note that there is a time lag to the SSN depending on the solar wind parameter. Also plotted are the radial dependencies of the analytic DQCS magnetic field model for solar minimum from \citet{Banaszkiewicz1998}, the slow wind velocity models from \citet{Wang2000} and \citet{Sheeley1997}, the density model from \citet{Leblanc1998} and the range of temperature measurements from \citet{Billings1959} and \citet{Liebenberg1975}.}
	\label{fig:sw_extrapolation_ssn_d_plot}
\end{figure*}

@@

%velocity:
The average velocity is found to decrease from \SI{340}{\km\per\s} at \SI{0.25}{\au} to about \SI{290}{\km\per\s} at \SI{0.046}{\au} for a SSN of 0. Whereas using a SSN of 200 it decreases from \SI{390}{\km\per\s} to \SI{330}{\km\per\s}. Comparing the results with the measurements by \citet{Sheeley1997} and \citet{Wang2000} shows an overestimation in our extrapolated slow solar wind velocity values for distances below about \SI{20}{\Rs}. They used LASCO coronagraph observations to track moving coronal features (blobs) in the distance range \SIrange{2}{30}{\Rs} to determine speed profiles and sources of the slow solar wind and they derived temperature and sonic point values for slow solar wind with the isothermal expansion model from \citet{Parker1958}. Therefore, it generally can be expected that PSP will encounter a slower solar wind environment close to the Sun than our model estimates, thus PSP will measure solar wind acceleration processes \citep{McComas2008}, maybe even still at \SI{30}{\Rs} as the study by \citet{Sheeley1997} suggests.

@@

\section{Discussion}
\label{sec:discussion}
We started the development of the empirical solar wind environment model for the near-ecliptic PSP orbit by lognormally fitting the about 40~years of in situ near-Earth solar wind data collected in the OMNI database, using the frequency distributions of the key solar wind parameters magnetic field strength, velocity, density and temperature. The OMNI multi-spacecraft solar wind data is intercalibrated and covers almost five solar cycles. It thus represents solar wind gathered at different phases of solar activity in the ecliptic plane. In the next step we investigated the yearly variation of the solar wind distribution functions along with the SSN over 53~years and derived linear dependencies of the solar wind parameters with the SSN. The radial dependencies of the solar wind distribution functions were then analyzed using Helios~1 and 2 data for the distance range \SIrange{0.29}{0.98}{\au} in bins of \SI{0.01}{\au}. The herefrom derived power law fit functions were used to scale the prior calculated SSN-dependent \SI{1}{\au} distribution fit functions to the PSP orbit, combined with SSN predictions for the years 2018--2025, i.e., for the prime mission phase that gets PSP to closest perihelia of 9.86~solar radii distance. The reason for performing the analysis this way is based on the fact that the OMNI solar wind database is much larger than the Helios database. Though, it is clear that the calculated distribution functions just represent first order estimates of the real solar wind to be encountered by PSP. The solar wind environment to be encountered will depend at times of PSP on the structure of the solar corona and underlying photospheric magnetic field and on the evolution and interaction of individual solar wind streams and superimposed CMEs and shocks.

However, the derived results are in good agreement with existing studies about near-Sun solar wind magnetic field strenghts and densities as shown in Sect.~\ref{sec:model_extrapolation_to_psp_orbit}. The from direct measurements differing velocity and temperature extrapolations below about 20~solar radii indicate that PSP will indeed dive into the acceleration and heating regions of the solar wind to be expected at these distances (see Fig.~\ref{fig:sw_extrapolation_ssn_d_plot}). The near-Sun solar wind velocity at PSP perihelion is also expected to be slower than our model estimates, because the region of the Alfvénic critical surface, up to which the solar wind is believed to be accelerated, is predicted to lie in average around \SI{17}{\Rs} \citep[e.g.,][]{Sittler1999,Exarhos2000} and its position correlates with solar activity between \SI{15}{\Rs} in cycle minimum and \SI{30}{\Rs} in maximum \citep{Katsikas2010,Goelzer2014}.	% in average around 17~Rs (e.g., 0.6~Rs Schatten 1969, 17~Rs Sittler1999, 17~Rs Exarhos2000) and scales with solar activity between 15~Rs in minimum and 30~Rs in maximum (19~Rs Katsikas2010, 15–30~Rs Goelzer2014).

Throughout the different analyses in our study the velocity's frequency distribution is treated as a composition of a slow and a fast wind distribution. Each velocity part is fitted with a lognormal function, which allows for the overlap of both velocity ranges.

We have not specifically investigated the occurrences of extreme solar wind parameters caused by CMEs or enhanced values due to stream interaction or co-rotating interaction regions. The Helios solar wind measurements plotted over radial distance in Fig.~\ref{fig:radial_fit_4_thesis_light_b_skip} show several extreme values far above the usual solar wind velocities, likely to be associated with individual CMEs. The results by \citet{Sachdeva2017} indicate that due to the prevailing of solar wind drag the speeds of fast CMEs will commonly slow down substantially from early distances of a few solar radii. Therefore, it is expected that PSP will encounter CMEs with much higher speeds than those observed during the Helios mission. Also, the magnetic field, density and temperature values are expected to be much larger than in the average solar wind in individual fast shock associated CME events. PSP will thus also substantially improve our understanding of the near-Sun evolution of CMEs and their expansion with radial distance.

%derive heliocentric distance depending error...\\
%error estimation for general model and extreme value tendencies\\

%magnetic field distribution's with distance increasing high value tail -> source are compression regions (why with density no increase?); look into Parker1958's B-field formula...\\
%varying shape with distance is indicator for internal physical processes (mixing/turbulence...)\\

%slow/fast streams basically maintain characteristic speeds with solar cycle; slow/fast ratio SSN dependency\\
%for the overlapping velocity model the SSN dependency is steeper than for a simple threshold\\
%The ratio of both varies with solar activity, e.g., 3~years after maximum, polar coronal holes are observed to often have equatorial extensions (cite?). see and use citet{Bougeret1984} p.~498...\\

\section{Summary}
\label{sec:summary}
For the near-Earth solar wind OMNI data and the Helios~1 and 2 data, obtained over the distance range \SIrange{0.29}{0.98}{\au}, we derived lognormal representations of the frequency distributions’ shapes of the four key solar wind parameters magnetic field strength, proton velocity, proton density and proton temperature. The dependencies of these frequency distributions on the solar activity cycle and on radial distance to the Sun have been modeled with analytical relations and extrapolated to the Parker Solar Probe orbit, taking into account predictions of the sunspot number. This empirical CGAUSS solar wind model for PSP, representing the solar wind’s solar activity and distance behavior, yields estimations for the solar wind median values at PSP’s first perihelion in 2018 at a solar distance of \SI{0.16}{\au} and at PSP’s closest perihelion in 2024 at \SI{0.046}{\au}. The model yields the following main results for the bulk solar wind parameters:
\begin{itemize}
	\item The empirical dependency of the magnetic field strength median value on solar activity and radial distance is
	\begin{align}
		B_\text{med}(ssn,r) = (\SI{0.0131}{\nT} \cdot ssn + \SI{4.29}{\nT}) \cdot r^{-1.66}	\,.
	\end{align}
	This approximation appears to be valid for PSP's full distance range, however near its closest perihelion the actual median values might lie within \SI{20}{\%} higher. The estimated median values for PSP's first and closest perihelion are \SI{87}{\nT} and \SI{943}{\nT}.
	\item The radial dependency of the median velocity value for slow solar wind is
	\begin{align}
		v_\text{slow}(r) &= \SI{363}{\km\per\s} \cdot r^{0.099}
		\intertext{and that for fast solar wind is}
		v_\text{fast}(r) &= \SI{483}{\km\per\s} \cdot r^{0.099}	\,.
	\end{align}
	These relations are valid above about \SI{20}{\Rs}, below this solar distance they overestimate the actual velocities obtained from remote measurements. Their combined frequency distribution (\ref{eq:double_lognormal_fit_function}) is balanced by $c(ssn) = -0.00180 \cdot ssn + 0.64$, that is, at solar minimum ($ssn = 0$) the slow wind has a share of \SI{64}{\%}, dropping to only \SI{28}{\%} during maximum ($ssn = 200$). The calculated median values for PSP's first and closest perihelion are \SI{340}{\km\per\s} and \SI{290}{\km\per\s}.
	\item The density relation is found to be
	\begin{align}
		n_\text{med}(ssn,r) = \left(\SI{0.0038}{\per\cm\cubed} \cdot ssn + \SI{4.50}{\per\cm\cubed}\right) \cdot r^{-2.11}	\,.
	\end{align}
	It seems valid throughout the full PSP orbital distance range, even down to about \SI{8}{\Rs}. The estimated median values for PSP's first and closest perihelion are \SI{4015}{\per\cm\cubed} and \SI{9733}{\per\cm\cubed}.
	\item The derived correlation function for the temperature is
	\begin{align}
		T_\text{med}(ssn,r) = (\SI{197}{\K} \cdot ssn + \SI{57300}{\K}) \cdot r^{-1.10}	\,.
	\end{align}
	The evaluation of the validity of this relation shows, that the temperature values near PSP's perihelion are too high in comparison to coronal measurements. The estimated median values for PSP's first and closest perihelion are \SI{503000}{\K} and \SI{1930000}{\K}.
\end{itemize}

The velocity and temperature values are overestimated below \SI{20}{\Rs} in comparison with existing observations. This shows that solar wind acceleration and heating processes below \SI{20}{\Rs} limit the simple back-extrapolation from existing in situ measurements.
